---
title: "Practical 7"
author: "Trevor Atkins"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```

#### a

Variables:

The independent variable is motivation as a nominal measure and the dependent variable is pronunciation (difference in vowel length) as a interval measure. The independent variable has 2 levels ("high" and "low").

Appropriate perspective: 

Comparing the means between the high and low motivation groups, the appropriate statistical test will be the t-test.

```{r}
data7a <- readRDS(file="data7a.rds")
str(data7a)
```

```{r}
by(data7a$difference, data7a$Motivation, mean)
by(data7a$difference, data7a$Motivation, min)
by(data7a$difference, data7a$Motivation, max)
by(data7a$difference, data7a$Motivation, sd)
```

**Table 1**: Descriptives for the two motivation groups

value | High Motivation | Low Motivation |
mean |    47           |      16.33      |
minimum |   -7         |      -11        |
maximum|    109        |      71         |
sd |     36.32         |      27.79      |


```{r}
library("pastecs")
by(data7a$difference, data7a$Motivation, stat.desc, basic=FALSE, norm=TRUE)
```

The Shapiro Wilk's p-value for the low motivation group is 0.00348802, which is less than the significance value of 0.05 so the low motivation group data is not normally distributed although the p-value for the high motivation group is 0.8987226 so it accepts the null hypothesis that the high motivation group data is normally distributed.

```{r}
hist(data7a$difference[data7a$Motivation =="high"], prob=TRUE, xlab = "difference", main = "histogram of difference for highly motivated students")
curve(dnorm(x, mean=mean(data7a$difference[data7a$Motivation =="high"]), sd=sd(data7a$difference[data7a$Motivation =="high"])), add=TRUE)
```
```{r}
hist(data7a$difference[data7a$Motivation =="low"], prob=TRUE, xlab = "difference", main = "histogram of difference for low motivated students")
curve(dnorm(x, mean=mean(data7a$difference[data7a$Motivation =="low"]), sd=sd(data7a$difference[data7a$Motivation =="low"])), add=TRUE)
```

**Figure 1: ** Boxplot of Pronunciation as a function of Motivation group

```{r}
boxplot(difference~Motivation, data=data7a, main="Pronunciation over Motivation", xlab="Motivation group", ylab="Difference in Vowel Length")
```

#### b


